### 2019-03-05 manuscript.tex
Cleaned up MBE template- to begin writing ASAP

\documentclass[nogrid]{MBE}%

\usepackage{url}

\jshort{mst}

\volname{}

\jvolume{0}

\jvol{}

\jissue{0}

\pubyear{2019}

\mstype{Article}

\artid{012}

\begin{document}

\title{Mode and tempo of promoter evolution in the Paramecium aurelia species complex}

\author[Raborn et al.]{R. Taylor \surname{Raborn},$^{\ast,1,2}$ Timothy Licknack,$^{1,2}$ Wanfeng Guo,$^{1,2}$ Shannon N. Snyder,$^{1,2}$ and Michael Lynch$^{1,2}$}

\address{$^{1}$Biodesign Center for Mechanisms of Evolution\\
$^{2}$School of Life Sciences, Arizona State University, Room C445 797 E. Tyler Street, Tempe, AZ 85281}

%\history{Received xx xxxxx 2019; reviews returned xx xxxx 2019; accepted xx xxxxx 2019}

\coresp{E-mail: rtraborn@asu.edu}

%\datade{The POPRES data were obtained from dbGaP (accession no. phs000145.v1.p1).}

\abstract{Goes here
}

\keyword{cis-regulatory regions, Paramecium, promoter evolution, TSS profiling.}

\maketitle

\section{{Introduction}\label{sec:Intro}}

Enter introduction here

\section{Demographic structure}

More words here

\subsection{Subsection 1}

\subsubsection{Subsection 2}

\paragraph{Paragraph header}

\begin{arabiclist}
\item Item 1

\item Item 2

\item Item 3
\end{arabiclist}

\begin{itemize}
\item Consider a fall in population induced by a decline in the number of births in the economy,
taking as given mortality and migration.

\item It is well known that a lower population growth raises the capital--labor ratio in the Solow--Swan
growth model.

\item The same property holds in Diamond's (1965) overlapping generations model, and it enhances welfare
as long as the economy is dynamically efficient; i.e., when the interest rate exceeds the
population growth rate.
\end{itemize}
 A similar trend is observed in the United
States and advanced European countries (Gustafsson and Kalwij, 2006), and also in Canada,
Australia, and New Zealand (Sardon, 2006). Interestingly, as pointed out by Bongaarts and Feeney
(1998), even when the cohort's lifetime fertility rate (the number of children a mother has in her
lifetime) does not fall, the delayed childbearing alone leads to a decline in the number of
childbirths, measured by the total period fertility rates (TPFRs). %Ogawa and Retherford (1993),
%Kohler et al. (2002), and Sobotka (2004) confirmed that, to a certain extent, the delay of
%marriage and motherhood is responsible for the observed period fertility rate decline (now known
%as the `tempo effect').

\section{Model\label{sec:Model}}

\subsection{Demographic structure}

i.e.:
\begin{equation}
\lambda_{t}=\left\{
\begin{array}
[c]{cc}%
0, & t<0,\\
\lambda, & t\geq0.
\end{array}
\right.  \label{eq:lambda}%
\end{equation}

\begin{figure}[t]
\begin{center}
\includegraphics[height=0.21\textheight]{flrf1.eps}
\end{center}
\caption{Fluctuations in Cohort Size $N_{t}$ over Generations.}%
\label{fig:popdynamics}%
\end{figure}

%\TWOfig{\includegraphics[width=10pc]{flrf1.eps}}{Fluctuations in
%Cohort Size $N_{t}$ over
%Generations.\label{fig:popdynamics}}{\includegraphics[width=10pc]{flrf1.eps}}{Dynamics
%of Labor Force $L_{t}$.\newline
%\phantom{\hskip5pc}\label{fig:labordynamics}}

where $C$ is a constant term defined as $C\equiv\beta\log\beta-(1+\beta
)\log(1+\beta)+\beta\log A\alpha+(1+\beta)\log A\left(  1-\alpha\right)  $.
Similarly, long-term welfare in the benchmark economy ($\lambda=0$) can be
written as:
\begin{equation}
U^{\ast}=\left(  1+\beta\right)  \log[A\alpha\left(  k^{\ast}\right)
^{2\alpha-1}+\left(  k^{\ast}\right)  ^{\alpha}]-\beta\left(  1-\alpha\right)
\log k^{\ast}+C. \label{eq:U_benchmark}%
\end{equation}

\begin{table}[!t]%1
\tableparts{\caption{SH test results on nuclear and mitochondrial phylogenetic trees}\label{tab1}}
{\begin{tabular*}{\columnwidth}{@{\extracolsep{\fill}}lld{6,0}d{6,0}@{}}\toprule Sequence data &
\mcc{Tree} & \mcc{$-\ln~L$} & \mcc{SH test $P$-value} \\\colrule mtDNA& mtDNA& -109219.5& 0.5 \\
[0.1pt]
mtDNA& Nuclear& -61720.8& \mcc{\hspace*{5pt}$<{0.00001}$} \\
Nuclear& mtDNA& -113033.1& \mcc{\hspace*{5pt}$<{0.00001}$} \\
Nuclear& Nuclear& -60699.9& 0.5 \\\botrule
\end{tabular*}}
{}
\end{table}

%
\section{Supplementary Material}
Supplementary tables S1?S7 and figures S1?S11 are available  at Molecular Biology and Evolution
online (http://www.mbe.oxfordjournals.org/).

\section{Acknowledgments}

The authors gratefully acknowledge the help of Robert Policastro and Gabriel Zentner at Indiana University Bloomington for technical assistance and use of laboratory facilities.

\bibliographystyle{natbib}%%%%natbib.sty
\bibliography{refs}%%%refs.bib

\end{document}

### 2019-03-05 manuscript.tex
Minor updates
Added Loren ipsum text as placeholder to test formatting.

@@
 
-\title{Mode and tempo of promoter evolution in the Paramecium aurelia species complex}
+\title{Mode and tempo of promoter evolution measured in the Paramecium aurelia species complex}
 
@@
 \address{$^{1}$Biodesign Center for Mechanisms of Evolution\\
-$^{2}$School of Life Sciences, Arizona State University, Room C445 797 E. Tyler Street, Tempe, AZ 85281}
+$^{2}$School of Life Sciences\\
+Arizona State University, Office C445 797 E. Tyler Street, Tempe, AZ 85281}
 
@@
 
-\abstract{Goes here
+\abstract{
+Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
 }
 
-\keyword{cis-regulatory regions, Paramecium, promoter evolution, TSS profiling.}
+\keyword{\textit{cis}-regulatory regions, \textit{Paramecium}, paralogs, promoter evolution, TSS profiling.}
 
@@
 
-Enter introduction here
+Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
 
@@
 
-More words here
+Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
 
@@
 
+Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+
 \subsubsection{Subsection 2}
 
+Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+
 \paragraph{Paragraph header}
+
+Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+
+\section{{Methods}\label{sec:Methods}}
+
+Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+
+\section{{Results}\label{sec:Results}}
+
+Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+
+\section{{Discussion}\label{sec:Discussion}}
+
+Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
 

### 2019-03-08 manuscript.tex
Wrote beginnings of introduction.

--- 2019-03-05 manuscript.tex
@@
 
-\title{Mode and tempo of promoter evolution in the Paramecium aurelia species complex}
+\title{Mode and tempo of promoter evolution in the \textit{Paramecium aurelia} species complex}
 
@@
 
-%\datade{The POPRES data were obtained from dbGaP (accession no. phs000145.v1.p1).}
-
 \abstract{Goes here
@@
 
-\keyword{cis-regulatory regions, Paramecium, promoter evolution, TSS profiling.}
+\keyword{cis-regulatory regions, duplicate genes, paralogous genes, Paramecium, promoter evolution, TSS profiling.}
 
@@
 
-Enter introduction here
+Genomic analysis across diverse eukaryotes has drawn attention to the important role of \textit{cis}-regulatory sequences in the evolution of gene expression \citep{Wittkopp:2008ki, Wittkopp:2011bc, Siepel:2014hd}.
+However, the \textit{cis}-regulatory sequence differences that accompany--or indeed underpin--species divergence remain largely unknown.
+A likely explanation for this is the difficulty in predicting \textit{cis}-regulatory regions from genomic sequence alone.
+Accurate estimation of promoter positions requires empirical, usually functional genomic evidence.
+At present, the best high-throughput approach to identify promoters at genome-scale is Transcription Start Site (TSS) profiling, which includes CAGE and RAMPAGE, among others.
+While TSS profiling methods differ methadologically, these methods capture the 5`-ends of capped mRNAs, sequence their corresponding cDNAs and align the reads to the genome to identify the TSSs present within a given transcriptome.
+Clustering of gene-adjacent TSSs defines a promoter, or transcription start region (TSR).
 

### 2019-03-08 manuscript.tex
Updates from Overleaf

@@
 
-\title{Mode and tempo of promoter evolution in the \textit{Paramecium aurelia} species complex}
+\title{Mode and tempo of promoter evolution as measured in the Paramecium aurelia species complex}
 

### 2019-03-08 manuscript.tex
Added italics to title.
New sentence to intro.

@@
 
-\title{Mode and tempo of promoter evolution as measured in the Paramecium aurelia species complex}
+\title{Mode and tempo of promoter evolution as measured in the \textit{Paramecium aurelia} species complex}
 
@@
 
-\address{$^{1}$Biodesign Center for Mechanisms of Evolution\\
+\address{$^{1}$Biodesign Center for the Mechanisms of Evolution\\
 $^{2}$School of Life Sciences\\
-Arizona State University, Office C445 797 E. Tyler Street, Tempe, AZ 85281}
+Arizona State University, 797 E. Tyler Street, Tempe, AZ 85281}
 
@@
 However, the \textit{cis}-regulatory sequence differences that accompany--or indeed underpin--species divergence remain largely unknown.
-A likely explanation for this is the difficulty in predicting \textit{cis}-regulatory regions from genomic sequence alone.
+One plausible explanation for this is the difficulty in predicting \textit{cis}-regulatory regions from genomic sequence alone.
 Accurate estimation of promoter positions requires empirical, usually functional genomic evidence.
 At present, the best high-throughput approach to identify promoters at genome-scale is Transcription Start Site (TSS) profiling, which includes CAGE and RAMPAGE, among others.
-While TSS profiling methods differ methadologically, these methods capture the 5`-ends of capped mRNAs, sequence their corresponding cDNAs and align the reads to the genome to identify the TSSs present within a given transcriptome.
-Clustering of gene-adjacent TSSs defines a promoter, or transcription start region (TSR).
+While TSS profiling methods differ among themselves methadologically, these methods capture the 5`-ends of capped mRNAs, sequence their corresponding cDNAs and align the reads to the genome to identify the TSSs present within a given transcriptome.
+Clustering of gene-adjacent TSSs defines a promoter, transcription start region (TSR), at single base-pair resolution, thereby providing genomic locations for \textit{cis}-regulatory regions within a species.\\
 

### 2019-03-12 manuscript.tex
Added material to Introduction.

@@
 
-Genomic analysis across diverse eukaryotes has drawn attention to the important role of \textit{cis}-regulatory sequences in the evolution of gene expression \citep{Wittkopp:2008ki, Wittkopp:2011bc, Siepel:2014hd}.
-However, the \textit{cis}-regulatory sequence differences that accompany--or indeed underpin--species divergence remain largely unknown.
+Investigation across diverse eukaryotes has drawn attention to the important role of \textit{cis}-regulatory sequences in the evolution of gene expression \citep{Wittkopp:2008ki, Wittkopp:2011bc, Siepel:2014hd}.
+However, the \textit{cis}-regulatory sequence differences that accompany---or indeed underpin---species divergence remain largely unknown.
 One plausible explanation for this is the difficulty in predicting \textit{cis}-regulatory regions from genomic sequence alone.
 Accurate estimation of promoter positions requires empirical, usually functional genomic evidence.
-At present, the best high-throughput approach to identify promoters at genome-scale is Transcription Start Site (TSS) profiling, which includes CAGE and RAMPAGE, among others.
-While TSS profiling methods differ among themselves methadologically, these methods capture the 5`-ends of capped mRNAs, sequence their corresponding cDNAs and align the reads to the genome to identify the TSSs present within a given transcriptome.
-Clustering of gene-adjacent TSSs defines a promoter, transcription start region (TSR), at single base-pair resolution, thereby providing genomic locations for \textit{cis}-regulatory regions within a species.\\
+At present, the best high-throughput approach to identify promoters at genome-scale is Transcription Start Site (TSS) profiling, which includes CAGE \citep{Murata:gf} and RAMPAGE \citep{Batut:2013fu}, among others.
+While TSS profiling methods differ among themselves methadologically, these methods capture the 5\prime$-ends of capped mRNAs, sequence their corresponding cDNAs and align the reads to the genome to identify the TSSs present within a given transcriptome.
+Clustering gene-adjacent TSSs defines a promoter, transcription start region (TSR), at single base-pair resolution, thereby providing genomic locations for \textit{cis}-regulatory regions within a species \citep{Lenhard:2012ena,Rach:2009ct}}.
+TSS profling studies in a variety of model organisms has demonstrated that the shape of TSS distributions (\textit{i.e.} promoter shape) at promoters is related to the promoter class and correlates with the function of the associated gene \citep{Carninci:2006in, Rach:2009ct,Hoskins:2011io,Raborn:2016cr}.
+In addition, recent work indicates that promoter shape is itself a quantitative trait \citep{Schor:2017fw}, which has important implications for the understanding of the evolution of \textit{cis}-regulatory regions.
+\\
 

### 2019-03-12 manuscript.tex
Updates from Overleaf

@@
 At present, the best high-throughput approach to identify promoters at genome-scale is Transcription Start Site (TSS) profiling, which includes CAGE \citep{Murata:gf} and RAMPAGE \citep{Batut:2013fu}, among others.
-While TSS profiling methods differ among themselves methadologically, these methods capture the 5\prime$-ends of capped mRNAs, sequence their corresponding cDNAs and align the reads to the genome to identify the TSSs present within a given transcriptome.
-Clustering gene-adjacent TSSs defines a promoter, transcription start region (TSR), at single base-pair resolution, thereby providing genomic locations for \textit{cis}-regulatory regions within a species \citep{Lenhard:2012ena,Rach:2009ct}}.
+While TSS profiling methods differ among themselves methadologically, these methods capture the 5$`$-ends of capped mRNAs, sequence their corresponding cDNAs and align the reads to the genome to identify the TSSs present within a given transcriptome.
+Clustering gene-adjacent TSSs defines a promoter, transcription start region (TSR), at single base-pair resolution, thereby providing genomic locations for \textit{cis}-regulatory regions within a species \citep{Lenhard:2012ena,Rach:2009ct}.
 TSS profling studies in a variety of model organisms has demonstrated that the shape of TSS distributions (\textit{i.e.} promoter shape) at promoters is related to the promoter class and correlates with the function of the associated gene \citep{Carninci:2006in, Rach:2009ct,Hoskins:2011io,Raborn:2016cr}.

### 2019-03-12 manuscript.tex
Updates from Overleaf

@@
 
-\title{Mode and tempo of promoter evolution as measured in the \textit{Paramecium aurelia} species complex}
+\title{Mode and tempo of promoter evolution as observed in the \textit{Paramecium aurelia} species complex}
 
@@
 
-\address{$^{1}$Biodesign Center for the Mechanisms of Evolution\\
+\address{$^{1}$Biodesign Institute Center for the Mechanisms of Evolution\\
 $^{2}$School of Life Sciences\\
@@
 
-\keyword{cis-regulatory regions, duplicate genes, paralogous genes, Paramecium, promoter evolution, TSS profiling.}
+\keyword{\textit{cis}-regulatory regions, gene duplicates, paralogous genes, Paramecium, promoter evolution, TSS profiling.}
 

### 2019-03-12 manuscript.tex

--- 2019-03-12 manuscript.tex
@@
 
-\keyword{cis-regulatory regions, duplicate genes, paralogous genes, Paramecium, promoter evolution, TSS profiling.}
+\keyword{\textit{cis}-regulatory regions, duplicate genes, paralogous genes, \textit{Paramecium}, promoter evolution, TSS profiling.}
 
@@
 
-Investigation across diverse eukaryotes has drawn attention to the important role of \textit{cis}-regulatory sequences in the evolution of gene expression \citep{Wittkopp:2008ki, Wittkopp:2011bc, Siepel:2014hd}.
-However, the \textit{cis}-regulatory sequence differences that accompany---or indeed underpin---species divergence remain largely unknown.
+Investigations across diverse eukaryotes have drawn attention to the important role of \textit{cis}-regulatory sequences in the evolution of gene expression \citep{Wittkopp:2008ki, Wittkopp:2011bc, Siepel:2014hd}.
+Work over the previous two decades have emphasized the contribution of \textit{cis}-regulatory sequence differences toward changes in gene expression (insert ref to Wittkopp 2004).
+However, the \textit{cis}-regulatory sequence differences that accompany---or indeed underpin---species divergence remain largely obscure.
 One plausible explanation for this is the difficulty in predicting \textit{cis}-regulatory regions from genomic sequence alone.
 Accurate estimation of promoter positions requires empirical, usually functional genomic evidence.
-At present, the best high-throughput approach to identify promoters at genome-scale is Transcription Start Site (TSS) profiling, which includes CAGE \citep{Murata:gf} and RAMPAGE \citep{Batut:2013fu}, among others.
-While TSS profiling methods differ among themselves methadologically, these methods capture the 5$`$-ends of capped mRNAs, sequence their corresponding cDNAs and align the reads to the genome to identify the TSSs present within a given transcriptome.
+At present, the best high-throughput approach to identify promoters at genome-scale is Transcription Start Site (TSS) profiling, which includes CAGE \citep{Murata:2014gpb} and RAMPAGE \citep{Batut:2013fu}, among others.
+While TSS profiling methods differ in technical between themselves, these protocols all capture the 5$`$-ends of capped mRNAs, sequence their corresponding cDNAs and align the reads to the genome to identify the TSSs present within a given transcriptome.
 Clustering gene-adjacent TSSs defines a promoter, transcription start region (TSR), at single base-pair resolution, thereby providing genomic locations for \textit{cis}-regulatory regions within a species \citep{Lenhard:2012ena,Rach:2009ct}.
 TSS profling studies in a variety of model organisms has demonstrated that the shape of TSS distributions (\textit{i.e.} promoter shape) at promoters is related to the promoter class and correlates with the function of the associated gene \citep{Carninci:2006in, Rach:2009ct,Hoskins:2011io,Raborn:2016cr}.
-In addition, recent work indicates that promoter shape is itself a quantitative trait \citep{Schor:2017fw}, which has important implications for the understanding of the evolution of \textit{cis}-regulatory regions.
+In addition, recent work done in \textit{Drosophila melanogaster} indicates that promoter shape is itself a quantitative trait \citep{Schor:2017fw}, a finding that carries valuable implications for the understanding of the evolution of \textit{cis}-regulatory regions.
 \\

### 2019-03-13 manuscript.tex
Added additional outline points from meeting with TL.

@@
 
-Investigations across diverse eukaryotes have drawn attention to the important role of \textit{cis}-regulatory sequences in the evolution of gene expression \citep{Wittkopp:2008ki, Wittkopp:2011bc, Siepel:2014hd}.
+[Ed: add 2 sentences on importance of gene expression] Investigations across select eukaryotes have drawn attention to the important role of \textit{cis}-regulatory sequences in the evolution of gene expression \citep{Wittkopp:2008ki, Wittkopp:2011bc, Siepel:2014hd}.
 Work over the previous two decades have emphasized the contribution of \textit{cis}-regulatory sequence differences toward changes in gene expression (insert ref to Wittkopp 2004).
-However, the \textit{cis}-regulatory sequence differences that accompany---or indeed underpin---species divergence remain largely obscure.
+[Add sentence(s) on gene expression -> transcription initiation.] [Summarize contributions in this area e.g.
+from Wittkopp 2004.] However, the \textit{cis}-regulatory sequence differences that accompany---or indeed underpin---species divergence remain largely obscure.
 One plausible explanation for this is the difficulty in predicting \textit{cis}-regulatory regions from genomic sequence alone.
@@
 \\
+-Add a paragraph introducing the Paramecium system -> the experiment\\
+-Final sentence(s) of intro: what we found plus the novelty eg this is the first study of its kind to do x.
 

### 2019-03-13 manuscript.tex
Added (new) first two sentences to Introduction.

@@
 
-[Ed: add 2 sentences on importance of gene expression] Investigations across select eukaryotes have drawn attention to the important role of \textit{cis}-regulatory sequences in the evolution of gene expression \citep{Wittkopp:2008ki, Wittkopp:2011bc, Siepel:2014hd}.
+The regulation of gene expression underpins fundamental processes within the cell, such as growth, development, the maintentance of homeostasis and metabolism.
+Expression of a gene is controlled through proximal, \textit{cis}-regulatory regions that are bound by \textit{trans}-acting transcription factors.
+Investigations across select eukaryotes have drawn attention to the important role of \textit{cis}-regulatory sequences in the evolution of gene expression differences \citep{Wittkopp:2008ki, Wittkopp:2011bc, Siepel:2014hd}.
 Work over the previous two decades have emphasized the contribution of \textit{cis}-regulatory sequence differences toward changes in gene expression (insert ref to Wittkopp 2004).

### 2019-03-13 manuscript.tex
Corrected spelling mistake

@@
 
-The regulation of gene expression underpins fundamental processes within the cell, such as growth, development, the maintentance of homeostasis and metabolism.
+The regulation of gene expression underpins fundamental processes within the cell, such as growth, development, the maintenance of homeostasis and metabolism.
 Expression of a gene is controlled through proximal, \textit{cis}-regulatory regions that are bound by \textit{trans}-acting transcription factors.

### 2019-03-13 manuscript.tex
Added reference to Wittkopp 2004 in intro and updated outline.

@@
 The regulation of gene expression underpins fundamental processes within the cell, such as growth, development, the maintenance of homeostasis and metabolism.
-Expression of a gene is controlled through proximal, \textit{cis}-regulatory regions that are bound by \textit{trans}-acting transcription factors.
+Expression of a gene is controlled through proximal, \textit{cis}-regulatory regions bound by \textit{trans}-acting transcription factors.
 Investigations across select eukaryotes have drawn attention to the important role of \textit{cis}-regulatory sequences in the evolution of gene expression differences \citep{Wittkopp:2008ki, Wittkopp:2011bc, Siepel:2014hd}.
-Work over the previous two decades have emphasized the contribution of \textit{cis}-regulatory sequence differences toward changes in gene expression (insert ref to Wittkopp 2004).
-[Add sentence(s) on gene expression -> transcription initiation.] [Summarize contributions in this area e.g.
+Work over the previous two decades have emphasized the contribution of \textit{cis}-regulatory sequence differences toward changes in gene expression \citep{Wittkopp:2004cy}.[Add sentence(s) on gene expression -> transcription initiation.] [Summarize contributions in this area e.g.
 from Wittkopp 2004.] However, the \textit{cis}-regulatory sequence differences that accompany---or indeed underpin---species divergence remain largely obscure.

### 2019-03-16 manuscript.tex
Added new first sentence to Intro.

@@
 
+A major challenge in genomics is to decipher the precise instructions that regulation gene expression.
 The regulation of gene expression underpins fundamental processes within the cell, such as growth, development, the maintenance of homeostasis and metabolism.
-Expression of a gene is controlled through proximal, \textit{cis}-regulatory regions bound by \textit{trans}-acting transcription factors.
+Transcription, a major step in gene expression, is controlled through proximal, \textit{cis}-regulatory regions bound by \textit{trans}-acting transcription factors.
 Investigations across select eukaryotes have drawn attention to the important role of \textit{cis}-regulatory sequences in the evolution of gene expression differences \citep{Wittkopp:2008ki, Wittkopp:2011bc, Siepel:2014hd}.

### 2019-03-24 manuscript.tex
Added more to introduction re gene expression.

@@
 Investigations across select eukaryotes have drawn attention to the important role of \textit{cis}-regulatory sequences in the evolution of gene expression differences \citep{Wittkopp:2008ki, Wittkopp:2011bc, Siepel:2014hd}.
-Work over the previous two decades have emphasized the contribution of \textit{cis}-regulatory sequence differences toward changes in gene expression \citep{Wittkopp:2004cy}.[Add sentence(s) on gene expression -> transcription initiation.] [Summarize contributions in this area e.g.
+Work over the previous two decades have emphasized the contribution of \textit{cis}-regulatory sequence differences toward changes in gene expression \citep{Wittkopp:2004cy}.
+The fulcrum and first major step of gene expression is transcription initiation, which in eukaryotes begins with the mediated interaction of RNA polymerase II complex with the promoter.
+[Summarize contributions in this area e.g.
 from Wittkopp 2004.] However, the \textit{cis}-regulatory sequence differences that accompany---or indeed underpin---species divergence remain largely obscure.

### 2019-03-27 manuscript.tex
More updates to intro.
Added new paragraph to Introduction.

@@
 
-A major challenge in genomics is to decipher the precise instructions that regulation gene expression.
+A major challenge in genomics is to decipher the precise instructions that regulate gene expression.
 The regulation of gene expression underpins fundamental processes within the cell, such as growth, development, the maintenance of homeostasis and metabolism.
@@
 The fulcrum and first major step of gene expression is transcription initiation, which in eukaryotes begins with the mediated interaction of RNA polymerase II complex with the promoter.
-[Summarize contributions in this area e.g.
-from Wittkopp 2004.] However, the \textit{cis}-regulatory sequence differences that accompany---or indeed underpin---species divergence remain largely obscure.
+The locations of eukaryotic \textit{cis}-regulatory sequences, including promoters, generally cannot be predicted from genomes with precision.\\
+Comparative and functional genomic approaches have shed some light on the contributions of \textit{cis}-regulatory sequences to phenotypic evolution and species divergence.
+Investigating F1 hybrids of \textit{D.
+melanogaster} and \textit{D.
+simulans}, Wittkopp and colleagues \citep{Wittkopp:2004cy} showed that interspecific expression differences were primarily caused by changes acting in \textit{cis}, not \textit{trans}.
+Evolutionary changes in gene expression patterns have been noted in a number of instances, including in expression timing (\textit{i.e.} heterochrony) \citep{Wray:1989ku}, expression levels (\citep{Crawford:1999ic}), spatial differences \citep{Abzhanov:2000vb}, and sex-specific expression \citep{Kopp:2000bl}; reviewed in \citep{Wray:2003kn}.
+The structure and conservation of the regulatory regions of select genes, including the hox \citep{Kmita:2002hg} and pax \citep{Plaza:1999kc} gene complexes.
+The advent of large-scale functional genomics methods has identified sequence variants within regulatory regions that contribute to expression differences [cite Pickrell et al., 2010, Lappalainen et al., 2013][Add one or two more sentences].
+However, the precise \textit{cis}-regulatory sequence differences that accompany---or indeed underpin---species divergence remain largely obscure.
 One plausible explanation for this is the difficulty in predicting \textit{cis}-regulatory regions from genomic sequence alone.

### 2019-03-28 manuscript.tex
More updates to Introduction.

@@
 Work over the previous two decades have emphasized the contribution of \textit{cis}-regulatory sequence differences toward changes in gene expression \citep{Wittkopp:2004cy}.
-The fulcrum and first major step of gene expression is transcription initiation, which in eukaryotes begins with the mediated interaction of RNA polymerase II complex with the promoter.
-The locations of eukaryotic \textit{cis}-regulatory sequences, including promoters, generally cannot be predicted from genomes with precision.\\
-Comparative and functional genomic approaches have shed some light on the contributions of \textit{cis}-regulatory sequences to phenotypic evolution and species divergence.
+The fulcrum and first major step of gene expression is transcription initiation, which in eukaryotes begins with the mediated interaction of RNA polymerase II complex with the promoter\citep{Kadonaga:2011gz}.
+The locations of eukaryotic \textit{cis}-regulatory sequences, including promoters, cannot be predicted from genomes with precision.\\
+Comparative and functional genomic approaches have shed light on the contributions of \textit{cis}-regulatory sequences to phenotypic evolution and species divergence.
 Investigating F1 hybrids of \textit{D.
@@
 The structure and conservation of the regulatory regions of select genes, including the hox \citep{Kmita:2002hg} and pax \citep{Plaza:1999kc} gene complexes.
-The advent of large-scale functional genomics methods has identified sequence variants within regulatory regions that contribute to expression differences [cite Pickrell et al., 2010, Lappalainen et al., 2013][Add one or two more sentences].
-However, the precise \textit{cis}-regulatory sequence differences that accompany---or indeed underpin---species divergence remain largely obscure.
+The advent of large-scale functional genomics methods has identified sequence variants within regulatory regions that contribute to expression differences \citep{Lappalainen:2013el,Pickrell:2010bq, Schor:2017fw}; a number of these have been functionally characterized.
+These efforts have revealed the cellular mechanisms of transcriptome and regulatory variation, they have largely been confined to major model systems, especially human, and have not focused on exploring the changes to \textit{cis}-regulatory regions over evolutionary time-scales.
+As it stands, the precise \textit{cis}-regulatory sequence differences that accompany---or indeed underpin---species divergence remain largely obscure.
 One plausible explanation for this is the difficulty in predicting \textit{cis}-regulatory regions from genomic sequence alone.
 Accurate estimation of promoter positions requires empirical, usually functional genomic evidence.
-At present, the best high-throughput approach to identify promoters at genome-scale is Transcription Start Site (TSS) profiling, which includes CAGE \citep{Murata:2014gpb} and RAMPAGE \citep{Batut:2013fu}, among others.
+At present, the best high-throughput approach to identify promoters at genome-scale is transcription start site (TSS) profiling, which includes CAGE \citep{Murata:2014gpb} and RAMPAGE \citep{Batut:2013fu}, among others.
 While TSS profiling methods differ in technical between themselves, these protocols all capture the 5$`$-ends of capped mRNAs, sequence their corresponding cDNAs and align the reads to the genome to identify the TSSs present within a given transcriptome.

### 2019-04-13 manuscript.tex
Updated preliminary title
Added stashed manuscript.txt file

@@
 
-\title{Mode and tempo of promoter evolution as observed in the \textit{Paramecium aurelia} species complex}
+\title{Tempo and mode of promoter evolution as observed in the \textit{Paramecium aurelia} species complex}
 
@@
 The regulation of gene expression underpins fundamental processes within the cell, such as growth, development, the maintenance of homeostasis and metabolism.
-Transcription, a major step in gene expression, is controlled through proximal, \textit{cis}-regulatory regions bound by \textit{trans}-acting transcription factors.
+Transcription, a major step in gene expression, is controlled through \textit{cis}-regulatory regions bound by \textit{trans}-acting transcription factors.
 Investigations across select eukaryotes have drawn attention to the important role of \textit{cis}-regulatory sequences in the evolution of gene expression differences \citep{Wittkopp:2008ki, Wittkopp:2011bc, Siepel:2014hd}.
 Work over the previous two decades have emphasized the contribution of \textit{cis}-regulatory sequence differences toward changes in gene expression \citep{Wittkopp:2004cy}.
-The fulcrum and first major step of gene expression is transcription initiation, which in eukaryotes begins with the mediated interaction of RNA polymerase II complex with the promoter\citep{Kadonaga:2011gz}.
+The fulcrum and first major step of gene expression is transcription initiation, which in eukaryotes begins with the mediated interaction of RNA polymerase II complex with the promoter, a \textit{cis}-regulatory region located proximal to the gene \citep{Kadonaga:2011gz}.
 The locations of eukaryotic \textit{cis}-regulatory sequences, including promoters, cannot be predicted from genomes with precision.\\
@@
 Evolutionary changes in gene expression patterns have been noted in a number of instances, including in expression timing (\textit{i.e.} heterochrony) \citep{Wray:1989ku}, expression levels (\citep{Crawford:1999ic}), spatial differences \citep{Abzhanov:2000vb}, and sex-specific expression \citep{Kopp:2000bl}; reviewed in \citep{Wray:2003kn}.
-The structure and conservation of the regulatory regions of select genes, including the hox \citep{Kmita:2002hg} and pax \citep{Plaza:1999kc} gene complexes.
-The advent of large-scale functional genomics methods has identified sequence variants within regulatory regions that contribute to expression differences \citep{Lappalainen:2013el,Pickrell:2010bq, Schor:2017fw}; a number of these have been functionally characterized.
-These efforts have revealed the cellular mechanisms of transcriptome and regulatory variation, they have largely been confined to major model systems, especially human, and have not focused on exploring the changes to \textit{cis}-regulatory regions over evolutionary time-scales.
+The structure and conservation of the regulatory regions of select genes, including the hox \citep{Kmita:2002hg} and pax \citep{Plaza:1999kc} gene complexes, were established.
+More recently, the advent of large-scale functional genomics methods has identified sequence variants within regulatory regions that contribute to expression differences \citep{Lappalainen:2013el,Pickrell:2010bq, Schor:2017fw}; a number of these have been functionally characterized.
+Global functional genomics efforts have revealed the cellular mechanisms of transcriptome and regulatory variation, but have largely been confined to major model systems, especially human, and have not focused on exploring the changes to \textit{cis}-regulatory regions over evolutionary time-scales.
 As it stands, the precise \textit{cis}-regulatory sequence differences that accompany---or indeed underpin---species divergence remain largely obscure.
-One plausible explanation for this is the difficulty in predicting \textit{cis}-regulatory regions from genomic sequence alone.
-Accurate estimation of promoter positions requires empirical, usually functional genomic evidence.
+Another potential explanation for this is the difficulty in predicting \textit{cis}-regulatory regions from genomic sequence alone.
+Accurate estimation of promoter positions requires empirical evidence.
 At present, the best high-throughput approach to identify promoters at genome-scale is transcription start site (TSS) profiling, which includes CAGE \citep{Murata:2014gpb} and RAMPAGE \citep{Batut:2013fu}, among others.

### 2019-07-10 manuscript.tex
Small updates to introduction.
More material in intro.

@@
 The fulcrum and first major step of gene expression is transcription initiation, which in eukaryotes begins with the mediated interaction of RNA polymerase II complex with the promoter, a \textit{cis}-regulatory region located proximal to the gene \citep{Kadonaga:2011gz}.
-The locations of eukaryotic \textit{cis}-regulatory sequences, including promoters, cannot be predicted from genomes with precision.\\
+The locations of eukaryotic \textit{cis}-regulatory sequences, including promoters, cannot be predicted from genome sequence alone with precision.\\
 Comparative and functional genomic approaches have shed light on the contributions of \textit{cis}-regulatory sequences to phenotypic evolution and species divergence.
@@
 Evolutionary changes in gene expression patterns have been noted in a number of instances, including in expression timing (\textit{i.e.} heterochrony) \citep{Wray:1989ku}, expression levels (\citep{Crawford:1999ic}), spatial differences \citep{Abzhanov:2000vb}, and sex-specific expression \citep{Kopp:2000bl}; reviewed in \citep{Wray:2003kn}.
-The structure and conservation of the regulatory regions of select genes, including the hox \citep{Kmita:2002hg} and pax \citep{Plaza:1999kc} gene complexes, were established.
+The structure and conservation of the regulatory regions of select genes, including the \textit{hox} \citep{Kmita:2002hg} and \textit{pax} \citep{Plaza:1999kc} gene complexes, were established.
 More recently, the advent of large-scale functional genomics methods has identified sequence variants within regulatory regions that contribute to expression differences \citep{Lappalainen:2013el,Pickrell:2010bq, Schor:2017fw}; a number of these have been functionally characterized.
@@
 Another potential explanation for this is the difficulty in predicting \textit{cis}-regulatory regions from genomic sequence alone.
-Accurate estimation of promoter positions requires empirical evidence.
-At present, the best high-throughput approach to identify promoters at genome-scale is transcription start site (TSS) profiling, which includes CAGE \citep{Murata:2014gpb} and RAMPAGE \citep{Batut:2013fu}, among others.
+As such, accurate estimation of promoter positions requires direct, experimental evidence.
+At present, the most efficient approach to identify promoters at genome-scale is transcription start site (TSS) profiling, which includes CAGE (Cap Analysis of Gene Expression) \citep{Murata:2014gpb} and RAMPAGE \citep{Batut:2013fu}, among others.
 While TSS profiling methods differ in technical between themselves, these protocols all capture the 5$`$-ends of capped mRNAs, sequence their corresponding cDNAs and align the reads to the genome to identify the TSSs present within a given transcriptome.
@@
 TSS profling studies in a variety of model organisms has demonstrated that the shape of TSS distributions (\textit{i.e.} promoter shape) at promoters is related to the promoter class and correlates with the function of the associated gene \citep{Carninci:2006in, Rach:2009ct,Hoskins:2011io,Raborn:2016cr}.
-In addition, recent work done in \textit{Drosophila melanogaster} indicates that promoter shape is itself a quantitative trait \citep{Schor:2017fw}, a finding that carries valuable implications for the understanding of the evolution of \textit{cis}-regulatory regions.
+In addition, recent work done in \textit{Drosophila melanogaster} provides evidence that promoter shape is itself a quantitative trait \citep{Schor:2017fw}, a finding that carries valuable implications for the understanding of the evolution of \textit{cis}-regulatory regions.
 \\
+A limited number of studies have investigated the evolution of promoters using TSS profiling information.
+Frith et al.
+\citep{Frith:2006bb} compared TSSs between human and mouse tissues using the first generation of CAGE data.
+The authors observed ~1250 instances of shifting promoter usage within homologous sites as well as tissue-specific TSS differences between species.
+Frith and colleagues also found that, consistent with expectations, TSSs with high turnover exhibited less sequence conservation in its promoter region than those without.
+The authors propose a model of gradual shifting between TSSs via alternative initiation sites, as well as shifting of intiation along the sequence itself.
+\\
+Main et al. \citep{Main:2013dg}, used TSS profiling information generated in \textit{D. melanogaster} to identify putative orthologous TSSs in three other \textit{Drosophila} taxa.
 -Add a paragraph introducing the Paramecium system -> the experiment\\

### 2020-06-30 manuscript.tex
Added more material to the introduction.

@@
 
-%\history{Received xx xxxxx 2019; reviews returned xx xxxx 2019; accepted xx xxxxx 2019}
+%\history{Received xx xxxxx 2020; reviews returned xx xxxx 2020; accepted xx xxxxx 2020}
 
@@
 
-\abstract{Goes here
-}
+\abstract{Goes here}
 
@@
 \\
-Main et al. \citep{Main:2013dg}, used TSS profiling information generated in \textit{D. melanogaster} to identify putative orthologous TSSs in three other \textit{Drosophila} taxa.
+Main et al.
+\citep{Main:2013dg}, used a separate TSS profiling method in \textit{D.
+melanogaster} to identify putative orthologous TSSs in three other \textit{Drosophila} taxa: \textit{D.
+simulans}, \textit{D.
+pseudoobscura} and \textit{D.
+sechellia}.
+The authors found that the location and estimated activities of orthologous TSSs to be generally conserved, but did observe instances of TSS variation that suggest new promoters.
+Notably, Main \textit{Main et.
+al.} found species-specific differences in the alcohol dehydrogenase (ADH) gene among the four species.
 -Add a paragraph introducing the Paramecium system -> the experiment\\

### 2020-06-30 manuscript.tex
Additional writing in Introduction.

@@
 A major challenge in genomics is to decipher the precise instructions that regulate gene expression.
-The regulation of gene expression underpins fundamental processes within the cell, such as growth, development, the maintenance of homeostasis and metabolism.
-Transcription, a major step in gene expression, is controlled through \textit{cis}-regulatory regions bound by \textit{trans}-acting transcription factors.
-Investigations across select eukaryotes have drawn attention to the important role of \textit{cis}-regulatory sequences in the evolution of gene expression differences \citep{Wittkopp:2008ki, Wittkopp:2011bc, Siepel:2014hd}.
+Differences in gene regulatory regions regions has long been posited to underlie a large part of the diversification seen across species (xxxx King and Wilson, 1975), and indeed investigations across select eukaryotes have drawn attention to the contributions of \textit{cis}-regulatory sequences in the evolution of gene expression differences \citep{Wittkopp:2008ki, Wittkopp:2011bc, Siepel:2014hd}.
 Work over the previous two decades have emphasized the contribution of \textit{cis}-regulatory sequence differences toward changes in gene expression \citep{Wittkopp:2004cy}.
@@
 A limited number of studies have investigated the evolution of promoters using TSS profiling information.
-Frith et al.
+Frith \textit{et al}.
 \citep{Frith:2006bb} compared TSSs between human and mouse tissues using the first generation of CAGE data.
@@
 The authors found that the location and estimated activities of orthologous TSSs to be generally conserved, but did observe instances of TSS variation that suggest new promoters.
-Notably, Main \textit{Main et.
-al.} found species-specific differences in the alcohol dehydrogenase (ADH) gene among the four species.
+Notably, \textit{Main et.
+al.} found differences in promoter selection in the Alcohol dehydrogenase (ADH) gene across the four species.
+Working with high-resolution promoter data in human and mouse, Young and colleagues \citep{Young:2015xx} found that more than half (56\$) of promoters were turned over between the two species.
+#continue here
 -Add a paragraph introducing the Paramecium system -> the experiment\\

### 2020-07-01 manuscript.tex
Small update to latter part of intro

@@
 al.} found differences in promoter selection in the Alcohol dehydrogenase (ADH) gene across the four species.
-Working with high-resolution promoter data in human and mouse, Young and colleagues \citep{Young:2015xx} found that more than half (56\$) of promoters were turned over between the two species.
-#continue here
+Working with high-resolution promoter data in human and mouse, Young and colleagues \citep{Young:2015xx} found that more than half (56\$) of tissue-matched promoters were turned over between the two species.
+From and measurements of selective constraint the concluded that promoter turnover is the major source of promoter gain and loss in mammals.
+xxxx fix that sentence
 -Add a paragraph introducing the Paramecium system -> the experiment\\

### 2020-07-02 manuscript.tex
Added additional material in latter part of Introduction. Next introduce paralogs.

@@
 Working with high-resolution promoter data in human and mouse, Young and colleagues \citep{Young:2015xx} found that more than half (56\$) of tissue-matched promoters were turned over between the two species.
-From and measurements of selective constraint the concluded that promoter turnover is the major source of promoter gain and loss in mammals.
-xxxx fix that sentence
+Noting that most protein-coding genes in mammals \citep{FANTOMConsortiumandtheRIKENPMIandCLSTDGT:2014hz} have multiple promoters, the authors propose that promoter turnover is an important but overlooked means by which transcriptional modules diverge during evolution.
+Add bit about paralogs and how we can use them to help us understand how \textit{cis}-regulatory regions evolve.
+-> we haven't studied the fate of cis-regulatory regions in ortho-paralog pairs.
+then:
 -Add a paragraph introducing the Paramecium system -> the experiment\\

### 2020-07-07 manuscript.tex
Added a bit to introduction.

@@
 Noting that most protein-coding genes in mammals \citep{FANTOMConsortiumandtheRIKENPMIandCLSTDGT:2014hz} have multiple promoters, the authors propose that promoter turnover is an important but overlooked means by which transcriptional modules diverge during evolution.
-Add bit about paralogs and how we can use them to help us understand how \textit{cis}-regulatory regions evolve.
--> we haven't studied the fate of cis-regulatory regions in ortho-paralog pairs.
-then:
+Interestingly, while the fates of ortho-paralog pairs that arise through whole genome duplication (WGD) has been heavily studied (enter citations for Ohno, 1970; Wolfe and Shields, 1997, Force et al., 1999, others), the behavior of \textit{cis}-regulatory regions of duplicate gene pairs is less well-understood.
+(Add a bit about paralogs and how we can use them to help us understand how \textit{cis}-regulatory regions evolve.
+-> we haven't studied the fate of cis-regulatory regions in ortho-paralog pairs.) then:
 -Add a paragraph introducing the Paramecium system -> the experiment\\

### 2020-07-28 manuscript.tex
Added reference to Li et al (Genome Research)

@@
 
-\author[Raborn et al.]{R. Taylor \surname{Raborn},$^{\ast,1,2}$ Timothy Licknack,$^{1,2}$ Wanfeng Guo,$^{1,2}$ Shannon N. Snyder,$^{1,2}$ and Michael Lynch$^{1,2}$}
+\author[Raborn et al.]{R. Taylor \surname{Raborn},$^{\ast,1,2}$ Timothy Licknack,$^{1,2}$ Shannon N. Snyder,$^{1,2}$ Tia Swenty,$^{1,2}$ Wanfeng Guo,$^{1,2}$ and Michael Lynch$^{1,2}$}
 
@@
 Frith and colleagues also found that, consistent with expectations, TSSs with high turnover exhibited less sequence conservation in its promoter region than those without.
-The authors propose a model of gradual shifting between TSSs via alternative initiation sites, as well as shifting of intiation along the sequence itself.
+The authors propose a model of gradual shifting between TSSs via alternative initiation sites, as well as shifting of initiation along the sequence itself.
 \\
@@
 The authors found that the location and estimated activities of orthologous TSSs to be generally conserved, but did observe instances of TSS variation that suggest new promoters.
-Notably, \textit{Main et.
+Notably, Main \textit{et.
 al.} found differences in promoter selection in the Alcohol dehydrogenase (ADH) gene across the four species.
@@
 Noting that most protein-coding genes in mammals \citep{FANTOMConsortiumandtheRIKENPMIandCLSTDGT:2014hz} have multiple promoters, the authors propose that promoter turnover is an important but overlooked means by which transcriptional modules diverge during evolution.
+Li and colleagues (2018) investigated the fates of TSSs across  the primate lineage.
+They found that new TSSs frequently emerge in transposable elements and repetitive regions of DNA, and are more likely to emerge in areas near existing regulatory regions.
+While this work was important, showing potential routes of regulatory innovation, the study was somewhat limited by the fact it only used TSSs empirically defined in human, obscuring the potential new promoters that emerged in the other primate taxa in the study.
+//
 Interestingly, while the fates of ortho-paralog pairs that arise through whole genome duplication (WGD) has been heavily studied (enter citations for Ohno, 1970; Wolfe and Shields, 1997, Force et al., 1999, others), the behavior of \textit{cis}-regulatory regions of duplicate gene pairs is less well-understood.

### 2020-07-30 manuscript.tex
Minor changes to intro

@@
 Notably, Main \textit{et.
-al.} found differences in promoter selection in the Alcohol dehydrogenase (ADH) gene across the four species.
+al.} found differences in promoter selection in the Alcohol dehydrogenase (ADH) gene across the four species \citep{Main:2013dg}.
 Working with high-resolution promoter data in human and mouse, Young and colleagues \citep{Young:2015xx} found that more than half (56\$) of tissue-matched promoters were turned over between the two species.
@@
 They found that new TSSs frequently emerge in transposable elements and repetitive regions of DNA, and are more likely to emerge in areas near existing regulatory regions.
-While this work was important, showing potential routes of regulatory innovation, the study was somewhat limited by the fact it only used TSSs empirically defined in human, obscuring the potential new promoters that emerged in the other primate taxa in the study.
+While this work was very valuable, showing potential routes of regulatory innovation, the study was limited by the fact it only used empirically-defined TSSs in human (as those from other primates were not available), a fact that prevented the study from identifying novel promoters in the other primate taxa in the study.
 //
-Interestingly, while the fates of ortho-paralog pairs that arise through whole genome duplication (WGD) has been heavily studied (enter citations for Ohno, 1970; Wolfe and Shields, 1997, Force et al., 1999, others), the behavior of \textit{cis}-regulatory regions of duplicate gene pairs is less well-understood.
+Interestingly, while the fates of ortho-paralog pairs that arise through whole genome duplication (WGD) has been heavily studied (enter citations for \citep{Ohno:1970wm}; Wolfe and Shields, 1997, Force et al., 1999, others), the behavior of \textit{cis}-regulatory regions of duplicate gene pairs is less well-understood.
 (Add a bit about paralogs and how we can use them to help us understand how \textit{cis}-regulatory regions evolve.

### 2020-07-30 manuscript.tex
Small changes to phrasing in introduction

@@
 \citep{Frith:2006bb} compared TSSs between human and mouse tissues using the first generation of CAGE data.
-The authors observed ~1250 instances of shifting promoter usage within homologous sites as well as tissue-specific TSS differences between species.
+The authors observed ~1250 instances of shifting promoter usage within homologous sites, as well as tissue-specific TSS differences between species.
 Frith and colleagues also found that, consistent with expectations, TSSs with high turnover exhibited less sequence conservation in its promoter region than those without.
@@
 sechellia}.
-The authors found that the location and estimated activities of orthologous TSSs to be generally conserved, but did observe instances of TSS variation that suggest new promoters.
+The authors found that the location and estimated activities of orthologous TSSs to be generally conserved, but did observe instances of TSS variation indicative of new promoters.
 Notably, Main \textit{et.
-al.} found differences in promoter selection in the Alcohol dehydrogenase (ADH) gene across the four species \citep{Main:2013dg}.
-Working with high-resolution promoter data in human and mouse, Young and colleagues \citep{Young:2015xx} found that more than half (56\$) of tissue-matched promoters were turned over between the two species.
+al.} found differences in promoter selection in the Alcohol dehydrogenase (ADH) gene across the four \textit{Drosophila} species \citep{Main:2013dg}.
+Working with high-resolution promoter data in human and mouse, Young and colleagues \citep{Young:2015xx} found that more than half (56\$) of tissue-matched promoters were lost, or turned over, between the two species.
 Noting that most protein-coding genes in mammals \citep{FANTOMConsortiumandtheRIKENPMIandCLSTDGT:2014hz} have multiple promoters, the authors propose that promoter turnover is an important but overlooked means by which transcriptional modules diverge during evolution.
@@
 They found that new TSSs frequently emerge in transposable elements and repetitive regions of DNA, and are more likely to emerge in areas near existing regulatory regions.
-While this work was very valuable, showing potential routes of regulatory innovation, the study was limited by the fact it only used empirically-defined TSSs in human (as those from other primates were not available), a fact that prevented the study from identifying novel promoters in the other primate taxa in the study.
+While this work was valuable, demonstrating potential routes of regulatory innovation, the study was limited by the fact it only used empirically-defined TSSs in human (as those from other primates were not available), a fact that prevented the study from identifying novel promoters in the other primate taxa in the study.
 //

### 2020-07-30 manuscript.tex
More updates to introduction

@@
 The locations of eukaryotic \textit{cis}-regulatory sequences, including promoters, cannot be predicted from genome sequence alone with precision.\\
+
 Comparative and functional genomic approaches have shed light on the contributions of \textit{cis}-regulatory sequences to phenotypic evolution and species divergence.
@@
 \\
+
 A limited number of studies have investigated the evolution of promoters using TSS profiling information.
@@
 \\
+
 Main et al.
@@
 //
-Interestingly, while the fates of ortho-paralog pairs that arise through whole genome duplication (WGD) has been heavily studied (enter citations for \citep{Ohno:1970wm}; Wolfe and Shields, 1997, Force et al., 1999, others), the behavior of \textit{cis}-regulatory regions of duplicate gene pairs is less well-understood.
+
+Dpulicate genes are a major source of evolutionary change cite Moore and Puruggnan, 2006).
+The widespread availability of high-quality genome assemblies and genome annotations has allowed us to investigate duplicate gene evolution in a variety of phylogenetic contexts.
+However, while the fates of ortho-paralog pairs that arise through whole genome duplication (WGD) has been heavily studied (enter citations for \citep{Ohno:1970wm}; Wolfe and Shields, 1997, Force et al., 1999, others), the behavior of \textit{cis}-regulatory regions of duplicate gene pairs is less well-understood.
+
 (Add a bit about paralogs and how we can use them to help us understand how \textit{cis}-regulatory regions evolve.

### 2020-07-31 manuscript.tex
More additions to introduction

--- 2020-07-30 manuscript.tex
@@
 //
-Interestingly, while the fates of ortho-paralog pairs that arise through whole genome duplication (WGD) has been heavily studied (enter citations for \citep{Ohno:1970wm}; Wolfe and Shields, 1997, Force et al., 1999, others), the behavior of \textit{cis}-regulatory regions of duplicate gene pairs is less well-understood.
-(Add a bit about paralogs and how we can use them to help us understand how \textit{cis}-regulatory regions evolve.
--> we haven't studied the fate of cis-regulatory regions in ortho-paralog pairs.) then:
+
+Duplicate genes are a major source of evolutionary change (cite Moore and Puruggnan, 2006), and their evolutionary fates have long been a subject of intense study (cite several revent studies).
+The availability of high-quality genome assemblies and genome annotations has allowed us to investigate duplicate gene evolution in a variety of phylogenetic contexts (cite Gu et al., 2003 and others).
+However, while the fates of ortho-paralog pairs that arise through whole genome duplication (WGD) has been heavily studied (enter citations for \citep{Ohno:1970wm}; Wolfe and Shields, 1997, Force et al., 1999, others), the fates of \textit{cis}-regulatory regions of duplicate gene pairs is much less well-understood.
+(Discuss Haberer et al., Dong et al and Arsovski et al).
+
+(Add a bit about paralogs and how we can use them to help us understand how \textit{cis}-regulatory regions evolve. then:
 -Add a paragraph introducing the Paramecium system -> the experiment\\

### 2020-07-31 manuscript.tex

@@
 
-Duplicate genes are a major source of evolutionary change (cite Moore and Puruggnan, 2006), and their evolutionary fates have long been a subject of intense study (cite several revent studies).
-The availability of high-quality genome assemblies and genome annotations has allowed us to investigate duplicate gene evolution in a variety of phylogenetic contexts (cite Gu et al., 2003 and others).
+Duplicate genes are a major source of evolutionary change (cite \citep{Ohno:1970wm} Moore and Puruggnan, 2006), and their evolutionary fates have long been a subject of intense study (reviewed in  The availability of high-quality genome assemblies and genome annotations has allowed us to investigate duplicate gene evolution in a variety of phylogenetic contexts (cite Gu et al., 2003 and others).
 However, while the fates of ortho-paralog pairs that arise through whole genome duplication (WGD) has been heavily studied (enter citations for \citep{Ohno:1970wm}; Wolfe and Shields, 1997, Force et al., 1999, others), the fates of \textit{cis}-regulatory regions of duplicate gene pairs is much less well-understood.

### 2020-08-01 manuscript.tex
Significant expansion of the genome duplicates section of the intro; section is nearly complete.

--- 2020-07-31 manuscript.tex
@@
 While this work was valuable, demonstrating potential routes of regulatory innovation, the study was limited by the fact it only used empirically-defined TSSs in human (as those from other primates were not available), a fact that prevented the study from identifying novel promoters in the other primate taxa in the study.
-//
+\\
 
-Duplicate genes are a major source of evolutionary change (cite Moore and Puruggnan, 2006), and their evolutionary fates have long been a subject of intense study (cite several revent studies).
+Duplicate genes are a major source of evolutionary change (cite \citep{Ohno:1970wm} Moore and Puruggnan, 2006), and their evolutionary fates have long been a subject of intense study (reviewed in Taylor and Raes).
 The availability of high-quality genome assemblies and genome annotations has allowed us to investigate duplicate gene evolution in a variety of phylogenetic contexts (cite Gu et al., 2003 and others).
-However, while the fates of ortho-paralog pairs that arise through whole genome duplication (WGD) has been heavily studied (enter citations for \citep{Ohno:1970wm}; Wolfe and Shields, 1997, Force et al., 1999, others), the fates of \textit{cis}-regulatory regions of duplicate gene pairs is much less well-understood.
-(Discuss Haberer et al., Dong et al and Arsovski et al).
+However, while the fates of ortho-paralog pairs has been heavily studied (enter citations for \citep{Ohno:1970wm}; Wolfe and Shields, 1997, Force et al., 1999, others), the fates of \textit{cis}-regulatory regions of duplicate gene pairs is much less well-understood.
+Haberer \textit{et al}.
+(2004) analyzed the \textit{cis}-regulatory regions of both tandem- and and segmantally duplicated genes in \textit{Arabidopsis thaliana}.
+Using sequence alignments, Haberer and colleages found evidence for conservation in the promoters of both tandem and segmental duplicates, with the more recently diverged tandem duplicate promoters having higher sequence similarity and gene expression correlation.\\
 
-(Add a bit about paralogs and how we can use them to help us understand how \textit{cis}-regulatory regions evolve. then:
+Working in the budding yeast \textit{Saccharomyces cerevisiae}, Dong \textit{et al.} (2011) investigated the relative contributions of \textit{cis}- and \textit{trans}-variability on gene expression variation.
+Dong and colleagues reported that, while \textit{trans}-regulation explained a larger proportion of the gene expression variation, expression divergence between paralogs was better explained by \textit{cis} effects.
+Additionally, Dong \textit{et al}.
+found that the promoters of gene duplicates contained more genetic diversity than did those of 'singletons' \textit{i.e.} genes not arising from gene duplication, and that the former have accumulated significantly more sequence polymorphisms than their singleton counterparts.\\
+
+The relative behavior of duplicate gene promoters that arose through different mechanisms are of interest.
+Arsvoski \textit{et al}., also working in \textit{A.
+thaliana}, incorporated DNAse I footprinting data into their analysis, asking whether duplicates from whole-genome duplicates (WGD) exhibited differences in their network connectivity compared with those that arose through tandem duplication.
+(Continue here)
+
 -Add a paragraph introducing the Paramecium system -> the experiment\\
@@
 
-\section{Demographic structure}
+\section{Methods}
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+\subsection{Culturing of \textit{Paramecium} strains}
+-Tim to fill in this section
 
-\subsection{Subsection 1}
+\subsection{STRIPE-seq library preparation}
+-Taylor to fill in this section
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+\subsection{Computational analysis of sequenced STRIPE-seq libraries}
+-Taylor to complete this section
 
-\subsubsection{Subsection 2}
-
-Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+\subsection{Identification of ortho-paralogs across \texit{Paramecium} spp.}
+-Taylor to complete this section
 

### 2020-08-03 manuscript.tex
Added several missing references

@@
 Noting that most protein-coding genes in mammals \citep{FANTOMConsortiumandtheRIKENPMIandCLSTDGT:2014hz} have multiple promoters, the authors propose that promoter turnover is an important but overlooked means by which transcriptional modules diverge during evolution.
-Li and colleagues (2018) investigated the fates of TSSs across  the primate lineage.
+Li and colleagues (2018) \citep{Li.2018} investigated the fates of TSSs across the primate lineage.
 They found that new TSSs frequently emerge in transposable elements and repetitive regions of DNA, and are more likely to emerge in areas near existing regulatory regions.
@@
 
-Duplicate genes are a major source of evolutionary change (cite \citep{Ohno:1970wm} Moore and Puruggnan, 2006), and their evolutionary fates have long been a subject of intense study (reviewed in Taylor and Raes).
-The availability of high-quality genome assemblies and genome annotations has allowed us to investigate duplicate gene evolution in a variety of phylogenetic contexts (cite Gu et al., 2003 and others).
-However, while the fates of ortho-paralog pairs has been heavily studied (enter citations for \citep{Ohno:1970wm}; Wolfe and Shields, 1997, Force et al., 1999, others), the fates of \textit{cis}-regulatory regions of duplicate gene pairs is much less well-understood.
+Duplicate genes are a major source of evolutionary change (cite \citep{Ohno:1970wm} Moore and Puruggnan, 2006), and their evolutionary fates have long been a subject of intense study (reviewed in \citep{Taylor.2004}).
+The availability of high-quality genome assemblies and genome annotations has allowed us to investigate duplicate gene evolution in a variety of phylogenetic contexts  \citep{Gu.2003} (cite others).
+However, while the fates of ortho-paralog pairs has been heavily studied (enter citations for \citep{Ohno:1970wm}; \citep{Wolfe.1997, Force.1999}, the fates of \textit{cis}-regulatory regions of duplicate gene pairs is much less well-understood.
 Haberer \textit{et al}.
 (2004) analyzed the \textit{cis}-regulatory regions of both tandem- and and segmantally duplicated genes in \textit{Arabidopsis thaliana}.
-Using sequence alignments, Haberer and colleages found evidence for conservation in the promoters of both tandem and segmental duplicates, with the more recently diverged tandem duplicate promoters having higher sequence similarity and gene expression correlation.\\
+Using sequence alignments, Haberer and colleagues found evidence for conservation in the promoters of both tandem and segmental duplicates, with the more recently diverged tandem duplicate promoters having higher sequence similarity and gene expression correlation.\\
 
@@
 The relative behavior of duplicate gene promoters that arose through different mechanisms are of interest.
-Arsvoski \textit{et al}., also working in \textit{A.
+Arsvoski \textit{et al}.
+\citep{Arsovski.2015}, also working in \textit{A.
 thaliana}, incorporated DNAse I footprinting data into their analysis, asking whether duplicates from whole-genome duplicates (WGD) exhibited differences in their network connectivity compared with those that arose through tandem duplication.

### 2020-08-03 manuscript.tex
Replaced overwritten citations with correct current citations

@@
 A major challenge in genomics is to decipher the precise instructions that regulate gene expression.
-Differences in gene regulatory regions regions has long been posited to underlie a large part of the diversification seen across species (xxxx King and Wilson, 1975), and indeed investigations across select eukaryotes have drawn attention to the contributions of \textit{cis}-regulatory sequences in the evolution of gene expression differences \citep{Wittkopp:2008ki, Wittkopp:2011bc, Siepel:2014hd}.
-Work over the previous two decades have emphasized the contribution of \textit{cis}-regulatory sequence differences toward changes in gene expression \citep{Wittkopp:2004cy}.
-The fulcrum and first major step of gene expression is transcription initiation, which in eukaryotes begins with the mediated interaction of RNA polymerase II complex with the promoter, a \textit{cis}-regulatory region located proximal to the gene \citep{Kadonaga:2011gz}.
+Differences in gene regulatory regions regions has long been posited to underlie a large part of the diversification seen across species (xxxx King and Wilson, 1975), and indeed investigations across select eukaryotes have drawn attention to the contributions of \textit{cis}-regulatory sequences in the evolution of gene expression differences \citep{Wittkopp.2008, Wittkopp.2011, Siepel.2014}.
+Work over the previous two decades have emphasized the contribution of \textit{cis}-regulatory sequence differences toward changes in gene expression \citep{Wittkopp.2004}.
+The fulcrum and first major step of gene expression is transcription initiation, which in eukaryotes begins with the mediated interaction of RNA polymerase II complex with the promoter, a \textit{cis}-regulatory region located proximal to the gene \citep{Kadonaga.2012}.
 The locations of eukaryotic \textit{cis}-regulatory sequences, including promoters, cannot be predicted from genome sequence alone with precision.\\
@@
 melanogaster} and \textit{D.
-simulans}, Wittkopp and colleagues \citep{Wittkopp:2004cy} showed that interspecific expression differences were primarily caused by changes acting in \textit{cis}, not \textit{trans}.
-Evolutionary changes in gene expression patterns have been noted in a number of instances, including in expression timing (\textit{i.e.} heterochrony) \citep{Wray:1989ku}, expression levels (\citep{Crawford:1999ic}), spatial differences \citep{Abzhanov:2000vb}, and sex-specific expression \citep{Kopp:2000bl}; reviewed in \citep{Wray:2003kn}.
-The structure and conservation of the regulatory regions of select genes, including the \textit{hox} \citep{Kmita:2002hg} and \textit{pax} \citep{Plaza:1999kc} gene complexes, were established.
-More recently, the advent of large-scale functional genomics methods has identified sequence variants within regulatory regions that contribute to expression differences \citep{Lappalainen:2013el,Pickrell:2010bq, Schor:2017fw}; a number of these have been functionally characterized.
+simulans}, Wittkopp and colleagues \citep{Wittkopp.2004} showed that interspecific expression differences were primarily caused by changes acting in \textit{cis}, not \textit{trans}.
+Evolutionary changes in gene expression patterns have been noted in a number of instances, including in expression timing (\textit{i.e.} heterochrony) \citep{Wray.1989}, expression levels \citep{Crawford.1999}, spatial differences \citep{Abzhanov.2000}, and sex-specific expression \citep{Kopp.2000}; reviewed in \citep{Wray.2003}.
+The structure and conservation of the regulatory regions of select genes, including the \textit{hox} \citep{Kmita.2002} and \textit{pax} \citep{Plaza.1999} gene complexes, were established.
+More recently, the advent of large-scale functional genomics methods has identified sequence variants within regulatory regions that contribute to expression differences \citep{Lappalainen.2013, Pickrell.2010, Schor.2017}; a number of these have been functionally characterized.
 Global functional genomics efforts have revealed the cellular mechanisms of transcriptome and regulatory variation, but have largely been confined to major model systems, especially human, and have not focused on exploring the changes to \textit{cis}-regulatory regions over evolutionary time-scales.
@@
 As such, accurate estimation of promoter positions requires direct, experimental evidence.
-At present, the most efficient approach to identify promoters at genome-scale is transcription start site (TSS) profiling, which includes CAGE (Cap Analysis of Gene Expression) \citep{Murata:2014gpb} and RAMPAGE \citep{Batut:2013fu}, among others.
+At present, the most efficient approach to identify promoters at genome-scale is transcription start site (TSS) profiling, which includes CAGE (Cap Analysis of Gene Expression) \citep{Murata.2014} and RAMPAGE \citep{Batut.2013}, among others.
 While TSS profiling methods differ in technical between themselves, these protocols all capture the 5$`$-ends of capped mRNAs, sequence their corresponding cDNAs and align the reads to the genome to identify the TSSs present within a given transcriptome.
-Clustering gene-adjacent TSSs defines a promoter, transcription start region (TSR), at single base-pair resolution, thereby providing genomic locations for \textit{cis}-regulatory regions within a species \citep{Lenhard:2012ena,Rach:2009ct}.
-TSS profling studies in a variety of model organisms has demonstrated that the shape of TSS distributions (\textit{i.e.} promoter shape) at promoters is related to the promoter class and correlates with the function of the associated gene \citep{Carninci:2006in, Rach:2009ct,Hoskins:2011io,Raborn:2016cr}.
-In addition, recent work done in \textit{Drosophila melanogaster} provides evidence that promoter shape is itself a quantitative trait \citep{Schor:2017fw}, a finding that carries valuable implications for the understanding of the evolution of \textit{cis}-regulatory regions.
+Clustering gene-adjacent TSSs defines a promoter, transcription start region (TSR), at single base-pair resolution, thereby providing genomic locations for \textit{cis}-regulatory regions within a species \citep{Lenhard.2012, Rach.2009}.
+TSS profling studies in a variety of model organisms has demonstrated that the shape of TSS distributions (\textit{i.e.} promoter shape) at promoters is related to the promoter class and correlates with the function of the associated gene \citep{Carninci.2006, Rach.2009, Hoskins.2011, Raborn:2016cr}.
+In addition, recent work done in \textit{Drosophila melanogaster} provides evidence that promoter shape is itself a quantitative trait \citep{Schor.2017}, a finding that carries valuable implications for the understanding of the evolution of \textit{cis}-regulatory regions.
 \\
@@
 Frith \textit{et al}.
-\citep{Frith:2006bb} compared TSSs between human and mouse tissues using the first generation of CAGE data.
+\citep{Frith.2006} compared TSSs between human and mouse tissues using the first generation of CAGE data.
 The authors observed ~1250 instances of shifting promoter usage within homologous sites, as well as tissue-specific TSS differences between species.
@@
 Main et al.
-\citep{Main:2013dg}, used a separate TSS profiling method in \textit{D.
+\citep{Main.2013}, used a separate TSS profiling method in \textit{D.
 melanogaster} to identify putative orthologous TSSs in three other \textit{Drosophila} taxa: \textit{D.
@@
 Notably, Main \textit{et.
-al.} found differences in promoter selection in the Alcohol dehydrogenase (ADH) gene across the four \textit{Drosophila} species \citep{Main:2013dg}.
+al.} found differences in promoter selection in the Alcohol dehydrogenase (ADH) gene across the four \textit{Drosophila} species \citep{Main.2013}.
 Working with high-resolution promoter data in human and mouse, Young and colleagues \citep{Young:2015xx} found that more than half (56\$) of tissue-matched promoters were lost, or turned over, between the two species.
-Noting that most protein-coding genes in mammals \citep{FANTOMConsortiumandtheRIKENPMIandCLSTDGT:2014hz} have multiple promoters, the authors propose that promoter turnover is an important but overlooked means by which transcriptional modules diverge during evolution.
-Li and colleagues (2018) \citep{Li.2018} investigated the fates of TSSs across the primate lineage.
+Noting that most protein-coding genes in mammals \citep{Consortium.2014} have multiple promoters, the authors propose that promoter turnover is an important but overlooked means by which transcriptional modules diverge during evolution.
+Li and colleagues (2018) (add citation) investigated the fates of TSSs across the primate lineage.
 They found that new TSSs frequently emerge in transposable elements and repetitive regions of DNA, and are more likely to emerge in areas near existing regulatory regions.
@@
 
-Duplicate genes are a major source of evolutionary change (cite \citep{Ohno:1970wm} Moore and Puruggnan, 2006), and their evolutionary fates have long been a subject of intense study (reviewed in \citep{Taylor.2004}).
+Duplicate genes are a major source of evolutionary change (cite \citep{Ohno.1970} Moore and Puruggnan, 2006), and their evolutionary fates have long been a subject of intense study (reviewed in \citep{Taylor.2004}).
 The availability of high-quality genome assemblies and genome annotations has allowed us to investigate duplicate gene evolution in a variety of phylogenetic contexts  \citep{Gu.2003} (cite others).
-However, while the fates of ortho-paralog pairs has been heavily studied (enter citations for \citep{Ohno:1970wm}; \citep{Wolfe.1997, Force.1999}, the fates of \textit{cis}-regulatory regions of duplicate gene pairs is much less well-understood.
+However, while the fates of ortho-paralog pairs has been heavily studied (enter citations for \citep{Ohno.1970}; \citep{Wolfe.1997, Force.1999}, the fates of \textit{cis}-regulatory regions of duplicate gene pairs is much less well-understood.
 Haberer \textit{et al}.

### 2020-08-03 manuscript.tex
Corrected typo causing compile error and replaced outdated citations with their current citation version

@@
 A major challenge in genomics is to decipher the precise instructions that regulate gene expression.
-Differences in gene regulatory regions regions has long been posited to underlie a large part of the diversification seen across species (xxxx King and Wilson, 1975), and indeed investigations across select eukaryotes have drawn attention to the contributions of \textit{cis}-regulatory sequences in the evolution of gene expression differences \citep{Wittkopp.2008, Wittkopp.2011, Siepel.2014}.
+Differences in gene regulatory regions regions has long been posited to underlie a large part of the diversification seen across species \citep{King.1975}, and indeed investigations across select eukaryotes have drawn attention to the contributions of \textit{cis}-regulatory sequences in the evolution of gene expression differences \citep{Wittkopp.2008, Wittkopp.2011, Siepel.2014}.
 Work over the previous two decades have emphasized the contribution of \textit{cis}-regulatory sequence differences toward changes in gene expression \citep{Wittkopp.2004}.
@@
 Clustering gene-adjacent TSSs defines a promoter, transcription start region (TSR), at single base-pair resolution, thereby providing genomic locations for \textit{cis}-regulatory regions within a species \citep{Lenhard.2012, Rach.2009}.
-TSS profling studies in a variety of model organisms has demonstrated that the shape of TSS distributions (\textit{i.e.} promoter shape) at promoters is related to the promoter class and correlates with the function of the associated gene \citep{Carninci.2006, Rach.2009, Hoskins.2011, Raborn:2016cr}.
+TSS profling studies in a variety of model organisms has demonstrated that the shape of TSS distributions (\textit{i.e.} promoter shape) at promoters is related to the promoter class and correlates with the function of the associated gene \citep{Carninci.2006, Rach.2009, Hoskins.2011, Raborn.2016}.
 In addition, recent work done in \textit{Drosophila melanogaster} provides evidence that promoter shape is itself a quantitative trait \citep{Schor.2017}, a finding that carries valuable implications for the understanding of the evolution of \textit{cis}-regulatory regions.
@@
 al.} found differences in promoter selection in the Alcohol dehydrogenase (ADH) gene across the four \textit{Drosophila} species \citep{Main.2013}.
-Working with high-resolution promoter data in human and mouse, Young and colleagues \citep{Young:2015xx} found that more than half (56\$) of tissue-matched promoters were lost, or turned over, between the two species.
+Working with high-resolution promoter data in human and mouse, Young and colleagues \citep{Young.2015} found that more than half (56\$) of tissue-matched promoters were lost, or turned over, between the two species.
 Noting that most protein-coding genes in mammals \citep{Consortium.2014} have multiple promoters, the authors propose that promoter turnover is an important but overlooked means by which transcriptional modules diverge during evolution.
@@
 
-\subsection{Identification of ortho-paralogs across \texit{Paramecium} spp.}
+\subsection{Identification of ortho-paralogs across \textit{Paramecium} spp.}
 -Taylor to complete this section
@@
 Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
-
-\begin{arabiclist}
-\item Item 1
-
-\item Item 2
-
-\item Item 3
-\end{arabiclist}
-
-\begin{itemize}
-\item Consider a fall in population induced by a decline in the number of births in the economy,
-taking as given mortality and migration.
-
-\item It is well known that a lower population growth raises the capital--labor ratio in the Solow--Swan
-growth model.
-
-\item The same property holds in Diamond's (1965) overlapping generations model, and it enhances welfare
-as long as the economy is dynamically efficient; i.e., when the interest rate exceeds the
-population growth rate.
-\end{itemize}
- A similar trend is observed in the United
-States and advanced European countries (Gustafsson and Kalwij, 2006), and also in Canada,
-Australia, and New Zealand (Sardon, 2006). Interestingly, as pointed out by Bongaarts and Feeney
-(1998), even when the cohort's lifetime fertility rate (the number of children a mother has in her
-lifetime) does not fall, the delayed childbearing alone leads to a decline in the number of
-childbirths, measured by the total period fertility rates (TPFRs). %Ogawa and Retherford (1993),
-%Kohler et al. (2002), and Sobotka (2004) confirmed that, to a certain extent, the delay of
-%marriage and motherhood is responsible for the observed period fertility rate decline (now known
-%as the `tempo effect').
 
